---
title: "Mycobiome_DiversityMeasures"
author: "Catherine Purse"
date: "2024-08-15"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r packages, results='hide'}
x <- c("phyloseq","RColorBrewer", "tidyverse", "patchwork", "microViz", "microbiome", "MicEco", "DatawRappr", "nlme", "ggsignif", "vegan", "pairwiseAdonis")
lapply(x, require, character.only = TRUE)
```

# Colour palette
```{r}
cb_palette <- c("#ee8866", "#77aadd","#d3d3d3","#FDF28E","#8ED29B","#bbccee", "#AFC30A","#cc99cc","#C1E49C","#dd7788", "#009E73")
```

# Load data

* Load phyloseq object
```{r}
myco_10k_noout <- readRDS("myco_phyloseq.rds")
```

* Reorder Region and Age Group factor levels
```{r}
# Get sample data from the phyloseq object
myco_df <- as.data.frame(sample_data(myco_10k_noout))

# Reorder factor levels - Region
myco_df$type <- factor(myco_df$type, levels=c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon"))

# Rename Mid as Adult
myco_df$agegroup<-gsub("Mid","Adult",as.character(myco_df$agegroup))

# Reorder factor levels - Age Group
myco_df$agegroup <- factor(myco_df$agegroup, levels=c("Young", "Adult", "Aged"))

# Update the sample data in the phyloseq object
sample_data(myco_10k_noout) <- sample_data(myco_df)
```

* Agglomorate to species
```{r}
myco_sp <- tax_glom(myco_10k_noout, "Species")
```

# Beta - Ordination Plot
```{r}
custom_shapes <- c("Duodenum" = 16, "Jejunum" = 17, "Ileum" = 18, "Caecum" = 19, "ProxColon" = 15, "DistalColon" = 8)

set.seed(123)
mpa.ord <- ordinate(myco_sp, "NMDS", "bray")
mpa.ord.plot <- plot_ordination(myco_sp, mpa.ord, type="samples", color="agegroup", shape="type") +
  theme_bw() +
  scale_shape_manual(name = "Type", values = custom_shapes, labels = c("Duodenum", "Jejunum", "Ileum", "Caecum", "Proximal Colon", "Distal Colon")) +  #this enables me to alter labels in legend
  theme(text = element_text(size = 15)) +
  geom_point(size = 4) +
  scale_color_manual(name = "Age Group", values = cb_palette)
print(mpa.ord.plot)

# PERMANOVA
head(sample_data(myco_sp))
bray <- phyloseq::distance(myco_sp, method = "bray")
sam <- data.frame(sample_data(myco_sp))
adonis2(bray ~ agegroup + type + agegroup*type, data = sam, permutations = 999, strata = sam$name) #no significance

# Pairwise comparisons
pw.age <- pairwise.adonis2(bray ~ agegroup, data = sam, p_adjust_m = 'BH', strata = 'name')
pw.type <- pairwise.adonis2(bray ~ type, data = sam, p_adjust_m = 'BH', strata = 'agegroup')
```

* Get summary tables for the PERMANOVA pairwise comp.
```{r}
# Function to extract p-values from pairwise comparisons
extract_pvalues <- function(pw_result) {
  p_values <- sapply(pw_result[-1], function(x) x$`Pr(>F)`[1])
  names(p_values) <- gsub("_vs_", "-", names(p_values))
  return(p_values)
}

# Extract p-values
age_pvalues <- extract_pvalues(pw.age)
type_pvalues <- extract_pvalues(pw.type)

# Create summary tables
age_summary <- data.frame(
  Comparison = names(age_pvalues),
  p.value = age_pvalues,
  Significant = ifelse(age_pvalues < 0.05, "*", "")
)

type_summary <- data.frame(
  Comparison = names(type_pvalues),
  p.value = type_pvalues,
  Significant = ifelse(type_pvalues < 0.05, "*", "")
)

age_summary
type_summary

# Export
write.csv(age_summary, folder_path, row.names = TRUE)

write.csv(type_summary, folder_path, row.names = TRUE)
```

```{r}
tiff(file.path(folder_path, "mycobiome_ordination_plot.tiff"), units="in", width=6.5, height=5, res=300)
plot(mpa.ord.plot)
dev.off()
```

# Regions - Chao1
```{r}
chao <- boxplot_alpha(myco_sp,
                          index = "chao1",
                          x_var = "type",
                          )

p.chao <- chao + scale_x_discrete() +
  theme_bw() +
  labs(x="Spatial Region", y="Chao1") +
    theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position ='none') +
  scale_fill_manual(name = "Region", values = cb_palette)
p.chao
```

# Statistical comparisons - Chao1
```{r}
# Statistical comparisons - Chao1
chao.reg <- meta(myco_sp)
chao.reg$diversity <- microbiome::alpha(myco_sp, index = "all")$chao1

# Define the regions for releveling
regions <- c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon")

# Initialize an empty list to store results
chao_results_list <- list()

# Initialize a list to store all p-values
all_p_values <- c()

# Loop over each region, relevel, fit the model, and store the results
for (region in regions) {
  chao.reg$type <- relevel(chao.reg$type, ref = region)
  
  model <- lme(diversity ~ type, random = ~ 1 | name, data = chao.reg, method = "REML")
  
  st <- summary(model)
  tabl <- as.data.frame(st$tTable)  # Convert the tTable to a data frame
  
  # Add a column to identify the reference region
  tabl$Reference <- region
  
  # Add a column to indicate significance based on p-value
  tabl$Significant <- ifelse(tabl$`p-value` < 0.05, "Yes", "No")
  
  # Store the result in the list
  chao_results_list[[region]] <- tabl
  
    # Collect all p-values (excluding the intercept)
  all_p_values <- c(all_p_values, tabl$`p-value`[-1])
}

# Apply FDR correction to all p-values
adjusted_p_values <- p.adjust(all_p_values, method = "BH")

# Initialize an index to keep track of which adjusted p-value to use
adj_p_index <- 1

# Loop through the results again to add adjusted p-values and significance
for (region in regions) {
  tabl <- chao_results_list[[region]]
  
  # Add adjusted p-values (skip the intercept)
  tabl$adjusted_p_value <- c(NA, adjusted_p_values[adj_p_index:(adj_p_index + nrow(tabl) - 2)])
  
  # Update the index
  adj_p_index <- adj_p_index + nrow(tabl) - 1
  
  # Add a column to indicate significance based on adjusted p-value
  tabl$Significant <- ifelse(tabl$adjusted_p_value < 0.05, "Yes", "No")
  
  # Update the result in the list
  chao_results_list[[region]] <- tabl
}

# Combine all the results into a single data frame
chao_results_df <- do.call(rbind, chao_results_list)

# Display the combined data frame with significance column
chao_results_df

# Create the summary table
chao_summary_table <- chao_results_df %>%
  filter(grepl("^[A-Za-z]+\\.type[A-Za-z]+$", rownames(.))) %>%  # Filter rows with type comparisons
  mutate(
    Comparison = paste0(Reference, "-", sub(".*type", "", rownames(.))),
    Significant = ifelse(Significant == "Yes", "*", "")
  ) %>%
  select(Comparison, Value, `p-value`, Significant) %>%
  arrange(Comparison)

# Create a matrix of significance
chao_sig_matrix <- chao_summary_table %>%
  mutate(From = sub("-.*", "", Comparison),
         To = sub(".*-", "", Comparison)) %>%
  select(From, To, Significant) %>%
  pivot_wider(names_from = To, values_from = Significant, values_fill = "") %>%
  column_to_rownames("From")

# Make the matrix symmetrical and remove self-comparisons
for(i in rownames(chao_sig_matrix)) {
  for(j in colnames(chao_sig_matrix)) {
    if(i == j) {
      chao_sig_matrix[i, j] = ""
    } else if(chao_sig_matrix[i, j] == "" && chao_sig_matrix[j, i] != "") {
      chao_sig_matrix[i, j] = chao_sig_matrix[j, i]
    }
  }
}

# Reorder matrix
desired_order <- c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon")
chao_sig_matrix <- chao_sig_matrix %>%
  select(all_of(desired_order)) %>%
  slice(match(desired_order, rownames(.)))

# Print the results
print(chao_summary_table)
print(chao_sig_matrix)

# Export
write.csv(chao_summary_table, folder_path, row.names = TRUE)

as.data.frame(chao_sig_matrix)
write.csv(chao_sig_matrix, folder_path, row.names = TRUE)
```


# Regions - Inverse Simpson
```{r}
# Diversity Inverse Simpson
inverse_simpson <- boxplot_alpha(myco_sp,
                          index = "inverse_simpson",
                          x_var = "type",
                          )

p.inverse_simpson <- inverse_simpson + 
  scale_x_discrete() +
  theme_bw() +
  labs(x="Spatial Region", y="Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position ='none') +
  scale_fill_manual(name = "Region", values = cb_palette)
p.inverse_simpson
```

# Statistical comparisons - Diversity Inverse Simpson
```{r}
# Statistical comparisons - Inverse Simpson
inverse_simpson.reg <- meta(myco_sp)
inverse_simpson.reg$diversity <- microbiome::alpha(myco_sp, index = "all")$diversity_inverse_simpson

# Define the regions for releveling
regions <- c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon")

# Initialize an empty list to store results
inverse_simpson_results_list <- list()

# Initialize a list to store all p-values
all_p_values <- c()

# Loop over each region, relevel, fit the model, and store the results
for (region in regions) {
  inverse_simpson.reg$type <- relevel(inverse_simpson.reg$type, ref = region)
  
  model <- lme(diversity ~ type, random = ~ 1 | name, data = inverse_simpson.reg, method = "REML")
  
  st <- summary(model)
  tabl <- as.data.frame(st$tTable)  # Convert the tTable to a data frame
  
  # Add a column to identify the reference region
  tabl$Reference <- region
  
  # Add a column to indicate significance based on p-value
  tabl$Significant <- ifelse(tabl$`p-value` < 0.05, "Yes", "No")
  
  # Store the result in the list
  inverse_simpson_results_list[[region]] <- tabl
  
  # Collect all p-values (excluding the intercept)
  all_p_values <- c(all_p_values, tabl$`p-value`[-1])
}

# Apply FDR correction to all p-values
adjusted_p_values <- p.adjust(all_p_values, method = "BH")

# Initialize an index to keep track of which adjusted p-value to use
adj_p_index <- 1

# Loop through the results again to add adjusted p-values and significance
for (region in regions) {
  tabl <- inverse_simpson_results_list[[region]]
  
  # Add adjusted p-values (skip the intercept)
  tabl$adjusted_p_value <- c(NA, adjusted_p_values[adj_p_index:(adj_p_index + nrow(tabl) - 2)])
  
  # Update the index
  adj_p_index <- adj_p_index + nrow(tabl) - 1
  
  # Add a column to indicate significance based on adjusted p-value
  tabl$Significant <- ifelse(tabl$adjusted_p_value < 0.05, "Yes", "No")
  
  # Update the result in the list
  inverse_simpson_results_list[[region]] <- tabl
}

# Combine all the results into a single data frame
inverse_simpson_results_df <- do.call(rbind, inverse_simpson_results_list)

# Display the combined data frame with significance column
inverse_simpson_results_df

# Create the summary table
inverse_simpson_summary_table <- inverse_simpson_results_df %>%
  filter(grepl("^[A-Za-z]+\\.type[A-Za-z]+$", rownames(.))) %>%  # Filter rows with type comparisons
  mutate(
    Comparison = paste0(Reference, "-", sub(".*type", "", rownames(.))),
    Significant = ifelse(Significant == "Yes", "*", "")
  ) %>%
  select(Comparison, Value, `p-value`, Significant) %>%
  arrange(Comparison)

# Create a matrix of significance
inverse_simpson_sig_matrix <- inverse_simpson_summary_table %>%
  mutate(From = sub("-.*", "", Comparison),
         To = sub(".*-", "", Comparison)) %>%
  select(From, To, Significant) %>%
  pivot_wider(names_from = To, values_from = Significant, values_fill = "") %>%
  column_to_rownames("From")

# Make the matrix symmetrical and remove self-comparisons
for(i in rownames(inverse_simpson_sig_matrix)) {
  for(j in colnames(inverse_simpson_sig_matrix)) {
    if(i == j) {
      inverse_simpson_sig_matrix[i, j] = ""
    } else if(inverse_simpson_sig_matrix[i, j] == "" && inverse_simpson_sig_matrix[j, i] != "") {
      inverse_simpson_sig_matrix[i, j] = inverse_simpson_sig_matrix[j, i]
    }
  }
}

# Reorder matrix
desired_order <- c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon")
inverse_simpson_sig_matrix <- inverse_simpson_sig_matrix %>%
  select(all_of(desired_order)) %>%
  slice(match(desired_order, rownames(.)))

# Print the results
print(inverse_simpson_summary_table)
print(inverse_simpson_sig_matrix)

# Export
write.csv(inverse_simpson_summary_table, folder_path, row.names = TRUE)

as.data.frame(inverse_simpson_sig_matrix)
write.csv(inverse_simpson_sig_matrix, folder_path, row.names = TRUE)
```

# Regions - Shannon
```{r}
# Statistical comparisons - Shannon
diversity_shannon.reg <- meta(myco_sp)
diversity_shannon.reg$diversity <- microbiome::alpha(myco_sp, index = "all")$diversity_shannon

# Define the regions for releveling
regions <- c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon")

# Initialize an empty list to store results
diversity_shannon_results_list <- list()

# Initialize a list to store all p-values
all_p_values <- c()

# Loop over each region, relevel, fit the model, and store the results
for (region in regions) {
  diversity_shannon.reg$type <- relevel(diversity_shannon.reg$type, ref = region)
  
  model <- lme(diversity ~ type, random = ~ 1 | name, data = diversity_shannon.reg, method = "REML")
  
  st <- summary(model)
  tabl <- as.data.frame(st$tTable)  # Convert the tTable to a data frame
  
  # Add a column to identify the reference region
  tabl$Reference <- region
  
  # Add a column to indicate significance based on p-value
  tabl$Significant <- ifelse(tabl$`p-value` < 0.05, "Yes", "No")
  
  # Store the result in the list
  diversity_shannon_results_list[[region]] <- tabl
  
  # Collect all p-values (excluding the intercept)
  all_p_values <- c(all_p_values, tabl$`p-value`[-1])
}

# Apply FDR correction to all p-values
adjusted_p_values <- p.adjust(all_p_values, method = "BH")

# Initialize an index to keep track of which adjusted p-value to use
adj_p_index <- 1

# Loop through the results again to add adjusted p-values and significance
for (region in regions) {
  tabl <- diversity_shannon_results_list[[region]]
  
  # Add adjusted p-values (skip the intercept)
  tabl$adjusted_p_value <- c(NA, adjusted_p_values[adj_p_index:(adj_p_index + nrow(tabl) - 2)])
  
  # Update the index
  adj_p_index <- adj_p_index + nrow(tabl) - 1
  
  # Add a column to indicate significance based on adjusted p-value
  tabl$Significant <- ifelse(tabl$adjusted_p_value < 0.05, "Yes", "No")
  
  # Update the result in the list
  diversity_shannon_results_list[[region]] <- tabl
}

# Combine all the results into a single data frame
diversity_shannon_results_df <- do.call(rbind, diversity_shannon_results_list)

# Display the combined data frame with significance column
diversity_shannon_results_df

# Create the summary table
diversity_shannon_summary_table <- diversity_shannon_results_df %>%
  filter(grepl("^[A-Za-z]+\\.type[A-Za-z]+$", rownames(.))) %>%  # Filter rows with type comparisons
  mutate(
    Comparison = paste0(Reference, "-", sub(".*type", "", rownames(.))),
    Significant = ifelse(Significant == "Yes", "*", "")
  ) %>%
  select(Comparison, Value, `p-value`, Significant) %>%
  arrange(Comparison)

# Create a matrix of significance
diversity_shannon_sig_matrix <- diversity_shannon_summary_table %>%
  mutate(From = sub("-.*", "", Comparison),
         To = sub(".*-", "", Comparison)) %>%
  select(From, To, Significant) %>%
  pivot_wider(names_from = To, values_from = Significant, values_fill = "") %>%
  column_to_rownames("From")

# Make the matrix symmetrical and remove self-comparisons
for(i in rownames(diversity_shannon_sig_matrix)) {
  for(j in colnames(diversity_shannon_sig_matrix)) {
    if(i == j) {
      diversity_shannon_sig_matrix[i, j] = ""
    } else if(diversity_shannon_sig_matrix[i, j] == "" && diversity_shannon_sig_matrix[j, i] != "") {
      diversity_shannon_sig_matrix[i, j] = diversity_shannon_sig_matrix[j, i]
    }
  }
}

# Reorder matrix
desired_order <- c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon")
diversity_shannon_sig_matrix <- diversity_shannon_sig_matrix %>%
  select(all_of(desired_order)) %>%
  slice(match(desired_order, rownames(.)))

# Print the results
print(diversity_shannon_summary_table)
print(diversity_shannon_sig_matrix)

# Export
write.csv(diversity_shannon_summary_table, folder_path, row.names = TRUE)

as.data.frame(diversity_shannon_sig_matrix)
write.csv(diversity_shannon_sig_matrix, folder_path, row.names = TRUE)

```

# Regions - Export
* Export
```{r}
reg.alpha_patchwork <- p.chao / p.inverse_simpson / p.shannon +
   plot_layout(guides = "collect", axis_titles = "collect")
reg.alpha_patchwork

tiff(file.path(folder_path, "mycobiome_alpha_patchwork_regions.tiff"), units="in", width=8, height=6, res=300)
plot(reg.alpha_patchwork)
dev.off()
```

# Age - Chao1
```{r}
# Age - D
myco.D <- subset_samples(myco_sp, type=="Duodenum")

chao1.age.D <- boxplot_alpha(myco.D,
                          index = "chao1",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Chao1") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") +
  ggtitle("Duodenum")
chao1.age.D

ch.D <- meta(myco.D)
ch.D$diversity <- microbiome::alpha(myco.D, index = "all")$chao1
kruskal.test(diversity ~ agegroup, data = ch.D)


# Age - J
myco.J <- subset_samples(myco_sp, type=="Jejunum")

chao1.age.J <- boxplot_alpha(myco.J,
                          index = "chao1",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Chao1") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") +
  ggtitle("Jejunum")
chao1.age.J

ch.J <- meta(myco.J)
ch.J$diversity <- microbiome::alpha(myco.J, index = "all")$chao1
kruskal.test(diversity ~ agegroup, data = ch.J)

# Age - I
myco.I <- subset_samples(myco_sp, type=="Ileum")

chao1.age.I <- boxplot_alpha(myco.I,
                          index = "chao1",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Chao1") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") +
  ggtitle("Ileum")
chao1.age.I

ch.I <- meta(myco.I)
ch.I$diversity <- microbiome::alpha(myco.I, index = "all")$chao1
kruskal.test(diversity ~ agegroup, data = ch.I)

# Age - C
myco.C <- subset_samples(myco_sp, type=="Caecum")

chao1.age.C <- boxplot_alpha(myco.C,
                          index = "chao1",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Chao1") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") +
  ggtitle("Caecum")
chao1.age.C

ch.C <- meta(myco.C)
ch.C$diversity <- microbiome::alpha(myco.C, index = "all")$chao1
kruskal.test(diversity ~ agegroup, data = ch.C)

# Age - PC
myco.PC <- subset_samples(myco_sp, type=="ProxColon")

chao1.age.PC <- boxplot_alpha(myco.PC,
                          index = "chao1",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Chao1") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") +
  ggtitle("PC")
chao1.age.PC

ch.PC <- meta(myco.PC)
ch.PC$diversity <- microbiome::alpha(myco.PC, index = "all")$chao1
kruskal.test(diversity ~ agegroup, data = ch.PC) # p = 0.02
pairwise.wilcox.test(ch.PC$diversity, ch.PC$agegroup,
                 p.adjust.method = "BH") # does not find significant pairwise comparisons

# Age - DC
myco.DC <- subset_samples(myco_sp, type=="DistalColon")

chao1.age.DC <- boxplot_alpha(myco.DC,
                          index = "chao1",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Chao1") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") +
  ggtitle("DC")
chao1.age.DC

ch.DC <- meta(myco.DC)
ch.DC$diversity <- microbiome::alpha(myco.DC, index = "all")$chao1
kruskal.test(diversity ~ agegroup, data = ch.DC)
```

*Make dataframe 
```{r}

Chao_data <- data.frame(
  Test = rep("Kruskal-Wallis", 6),
  Region = c("D", "J", "I", "C", "PC", "DC"),
  Chi_squared = c(1.5497, 3.7778, 3.6964, 1.9341, 6.758, 4.3556),
  df = rep(2, 6),
  p_value = c(0.4608, 0.1512, 0.1575, 0.3802, 0.03408, 0.1133)
)

Chao_data <- rbind(Chao_data, 
                   data.frame(Test = "Wilcoxon", 
                              Region = c("PC (Young vs Adult)","PC (Young vs Aged)","PC (Adult vs Aged)"),
                              Chi_squared = rep(NA,3),
                              df = rep(NA,3),
                              p_value = c(0.14,0.14,0.14)
                              ))

print(Chao_data)

# Export
write.csv(Chao_data, folder_path, row.names = TRUE)
```

* Chao1 - Patchwork
```{r}
# Patchwork plot
chao1.patchwork <- chao1.age.D + chao1.age.J + chao1.age.I + chao1.age.C + chao1.age.PC + chao1.age.DC +
  plot_layout(guides = "collect", axis_titles = "collect", ncol = 6)
chao1.patchwork
```


# Age - Inverse Simpson
```{r}
# Age - D
myco.D <- subset_samples(myco_sp, type=="Duodenum")

diversity_inverse_simpson.age.D <- boxplot_alpha(myco.D,
                          index = "diversity_inverse_simpson",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_inverse_simpson.age.D

is.D <- meta(myco.D)
is.D$diversity <- microbiome::alpha(myco.D, index = "all")$diversity_inverse_simpson
kruskal.test(diversity ~ agegroup, data = is.D)

# Age - J
myco.J <- subset_samples(myco_sp, type=="Jejunum")

diversity_inverse_simpson.age.J <- boxplot_alpha(myco.J,
                          index = "diversity_inverse_simpson",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_inverse_simpson.age.J

is.J <- meta(myco.J)
is.J$diversity <- microbiome::alpha(myco.J, index = "all")$diversity_inverse_simpson
kruskal.test(diversity ~ agegroup, data = is.J)

# Age - I
myco.I <- subset_samples(myco_sp, type=="Ileum")

diversity_inverse_simpson.age.I <- boxplot_alpha(myco.I,
                          index = "diversity_inverse_simpson",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_inverse_simpson.age.I

is.I <- meta(myco.I)
is.I$diversity <- microbiome::alpha(myco.I, index = "all")$diversity_inverse_simpson
kruskal.test(diversity ~ agegroup, data = is.I)

# Age - C
myco.C <- subset_samples(myco_sp, type=="Caecum")

diversity_inverse_simpson.age.C <- boxplot_alpha(myco.C,
                          index = "diversity_inverse_simpson",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_inverse_simpson.age.C

is.C <- meta(myco.C)
is.C$diversity <- microbiome::alpha(myco.C, index = "all")$diversity_inverse_simpson
kruskal.test(diversity ~ agegroup, data = is.C)

# Age - PC
myco.PC <- subset_samples(myco_sp, type=="ProxColon")

diversity_inverse_simpson.age.PC <- boxplot_alpha(myco.PC,
                          index = "diversity_inverse_simpson",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_inverse_simpson.age.PC

is.PC <- meta(myco.PC)
is.PC$diversity <- microbiome::alpha(myco.PC, index = "all")$diversity_inverse_simpson
kruskal.test(diversity ~ agegroup, data = is.PC)

# Age - DC
myco.DC <- subset_samples(myco_sp, type=="DistalColon")

diversity_inverse_simpson.age.DC <- boxplot_alpha(myco.DC,
                          index = "diversity_inverse_simpson",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Inverse Simpson") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_inverse_simpson.age.DC

is.DC <- meta(myco.DC)
is.DC$diversity <- microbiome::alpha(myco.DC, index = "all")$diversity_inverse_simpson
kruskal.test(diversity ~ agegroup, data = is.DC)
```

```{r}
InvSimpson_data <- data.frame(
  Test = rep("Kruskal-Wallis", 6),
  Region = c("D", "J", "I", "C", "PC", "DC"),
  Chi_squared = c(2.125, 0.69444, 1.0714, 0.35165, 1.5152, 3.8),
  df = rep(2, 6),
  W_statistic = rep(NA, 6),
  p_value = c(0.3456, 0.7066, 0.5853, 0.8388, 0.4688, 0.1496)
)

# Display the result
print(InvSimpson_data)

# Export
write.csv(InvSimpson_data, folder_path, row.names = TRUE)
```

* Inverse Simpson - Patchwork
```{r}
inverse_simpson.patchwork <- diversity_inverse_simpson.age.D + 
                                       diversity_inverse_simpson.age.J + 
                                       diversity_inverse_simpson.age.I + 
                                       diversity_inverse_simpson.age.C + 
                                       diversity_inverse_simpson.age.PC + 
                                       diversity_inverse_simpson.age.DC +
  plot_layout(guides = "collect", axis_titles = "collect", ncol = 6)

inverse_simpson.patchwork
```


# Age - Shannon
```{r}
# Age - D
myco.D <- subset_samples(myco_sp, type=="Duodenum")

diversity_shannon.age.D <- boxplot_alpha(myco.D,
                          index = "diversity_shannon",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Shannon") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_shannon.age.D

is.D <- meta(myco.D)
is.D$diversity <- microbiome::alpha(myco.D, index = "all")$diversity_shannon
kruskal.test(diversity ~ agegroup, data = is.D)

# Age - J
myco.J <- subset_samples(myco_sp, type=="Jejunum")

diversity_shannon.age.J <- boxplot_alpha(myco.J,
                          index = "diversity_shannon",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Shannon") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_shannon.age.J

is.J <- meta(myco.J)
is.J$diversity <- microbiome::alpha(myco.J, index = "all")$diversity_shannon
kruskal.test(diversity ~ agegroup, data = is.J)

# Age - I
myco.I <- subset_samples(myco_sp, type=="Ileum")

diversity_shannon.age.I <- boxplot_alpha(myco.I,
                          index = "diversity_shannon",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Shannon") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_shannon.age.I

is.I <- meta(myco.I)
is.I$diversity <- microbiome::alpha(myco.I, index = "all")$diversity_shannon
kruskal.test(diversity ~ agegroup, data = is.I)

# Age - C
myco.C <- subset_samples(myco_sp, type=="Caecum")

diversity_shannon.age.C <- boxplot_alpha(myco.C,
                          index = "diversity_shannon",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Shannon") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_shannon.age.C

is.C <- meta(myco.C)
is.C$diversity <- microbiome::alpha(myco.C, index = "all")$diversity_shannon
kruskal.test(diversity ~ agegroup, data = is.C)

# Age - PC
myco.PC <- subset_samples(myco_sp, type=="ProxColon")

diversity_shannon.age.PC <- boxplot_alpha(myco.PC,
                          index = "diversity_shannon",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Shannon") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_shannon.age.PC

is.PC <- meta(myco.PC)
is.PC$diversity <- microbiome::alpha(myco.PC, index = "all")$diversity_shannon
kruskal.test(diversity ~ agegroup, data = is.PC)

# Age - DC
myco.DC <- subset_samples(myco_sp, type=="DistalColon")

diversity_shannon.age.DC <- boxplot_alpha(myco.DC,
                          index = "diversity_shannon",
                          x_var = "agegroup",
                          ) + 
  scale_fill_manual(values=c("#bbccee", "#AFC30A","#cc99cc")) +
  xlab("Age Group") +
  ylab("Shannon") +
  theme(axis.text.x = element_text(size=10), 
        axis.text.y = element_text(size=10),
        axis.title.x = element_text(size = 12, color = "black"),
        axis.title.y = element_text(size = 12, color = "black"),
        axis.title = element_text(size = 12, color = "black"),
        plot.title = element_text(face = "bold", size = 12),
        legend.position="none") 
diversity_shannon.age.DC

is.DC <- meta(myco.DC)
is.DC$diversity <- microbiome::alpha(myco.DC, index = "all")$diversity_shannon
kruskal.test(diversity ~ agegroup, data = is.DC)
```

* Inverse Simpson - Patchwork
```{r}
shannon.patchwork <- diversity_shannon.age.D + 
                                       diversity_shannon.age.J + 
                                       diversity_shannon.age.I + 
                                       diversity_shannon.age.C + 
                                       diversity_shannon.age.PC + 
                                       diversity_shannon.age.DC +
  plot_layout(guides = "collect", axis_titles = "collect", ncol = 6)

shannon.patchwork
```

* Summarise stats data
```{r}
Shannon_data <- data.frame(
  Test = rep("Kruskal-Wallis", 6),
  Region = c("D", "J", "I", "C", "PC", "DC"),
  Chi_squared = c(2.125, 0.69444, 1.0714, 0.19341, 1.4091, 2.9556),
  df = rep(2, 6),
  W_statistic = rep(NA, 6),
  p_value = c(0.3456, 0.7066, 0.5853, 0.9078, 0.4943, 0.2281)
)

# Display the result
print(Shannon_data)

# Export
write.csv(Shannon_data, folder_path, row.names = TRUE)
```

# Export age comparison plots
```{r}
alpha_patchwork <- chao1.patchwork / inverse_simpson.patchwork / shannon.patchwork +
   plot_layout(guides = "collect", axis_titles = "collect")
alpha_patchwork

tiff(file.path(folder_path, "mycobiome_alpha_patchwork.tiff"), units="in", width=9.5, height=10, res=300)
plot(alpha_patchwork)
dev.off()
```
