---
title: "Mycobiome_Candida"
author: "Catherine Purse"
date: "2024-08-14"
output: html_document
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Load libraries
```{r packages, results='hide'}
x <- c("phyloseq","RColorBrewer", "tidyverse", "patchwork", "microViz")
lapply(x, require, character.only = TRUE)
```

# Colour palette
```{r}
cb_palette <- c("#ee8866", "#77aadd","#d3d3d3","#FDF28E","#8ED29B","#bbccee", "#AFC30A","#cc99cc","#C1E49C","#dd7788", "#009E73")
```

# Load data

* Load ASV/Taxa/Meta tables
```{r}
ASV <- read.csv("OTU-table.csv")
tax <- read.csv("taxonomy-table.csv")
meta <- read.csv("Metadata-file.csv")
```

* Reformat ASV table (turn ASV column into rows) and strip white space
```{r}
rownames(ASV) <- paste0("ASV", 1:nrow(ASV))
ASV <- subset(ASV, select = -c(X.NAME))
View(ASV)

#strip white space
ASV <- ASV %>%
  mutate(across(where(is.character), ~str_trim(., side = "both")))
```

* Same for taxonomy table
```{r}
#do the same for taxonomy table (need rownames to match)
rownames(tax) <- paste0("ASV", 1:nrow(tax))
tax <- subset(tax, select = -c(X.TAXONOMY))
View(tax)

#strip white space
tax <- tax %>%
  mutate(across(where(is.character), ~str_trim(., side = "both")))

tax[] <- lapply(tax, function(x) if(is.character(x)) gsub("\\s+", " ", x))

taxa <- unique(tax$Species)
taxa
```

* Same for metadata
```{r}
meta_df <- meta %>%
  column_to_rownames("X.NAME")
View(meta_df)

#strip white space
meta_df <- meta_df %>%
  mutate(across(where(is.character), ~str_trim(., side = "both")))
View(meta_df)
```

## Create phyloseq object
* Create phyloseq object
```{r}
#turn df in to matrices (required for phyloseq)
asv_mat <- as.matrix(ASV)
tax_mat <- as.matrix(tax)

#transform into phyloseq objects
ASV_ps <- otu_table(asv_mat, taxa_are_rows = TRUE)
tax_ps <- tax_table(tax_mat)
samples <- sample_data(meta_df)

#combine!
myco <- phyloseq(ASV_ps, tax_ps, samples)
myco
```

* Prune samples <10000 reads (as recommended in QIB best practice)
```{r}
myco_10k <- prune_samples(sample_sums(myco) > 10000, myco)
myco_10k
```

* Clean data of outliers and unclassified ASVs (Kingdom Level)
```{r}
#remove NHP 17 (M497BDC) as ID presented with diarrhoea
myco_10k_noout <- subset_samples(myco_10k, sample_names(myco_10k) != "NHP17Caecum" & sample_names(myco_10k) != "NHP17PC" & sample_names(myco_10k) != "NHP17DC")
sample_names(myco_10k_noout)
myco_10k_noout

# Remove unidentified ASVs at Kingdom level
myco_10k_noout <- subset_taxa(myco_10k_noout, Kingdom != "unclassified_Root")

# Check unique Phylum
unique_phylum <- unique(tax_table(myco_10k_noout)[, 'Phylum'])

# Print or view the unique data
print(unique_phylum)
```

* Transform sample counts
```{r}
myco_10k_noout <- transform_sample_counts(myco_10k_noout, function(x) x/sum(x))
```

# Filter to only include Candida pathobionts
```{r}
myco.candida <- subset_taxa(myco_10k_noout, Species %in% c("Candida_tropicalis", "Candida_albicans", "Candida_parapsilosis"))
```

# Melt into df and improve aesthetics
```{r}
myco.candida.df <- psmelt(myco.candida)
trimws(myco.candida.df)

# replace _ with " "
myco.candida.df$Species<-gsub("_"," ",as.character(myco.candida.df$Species))
```

* Plot
Some of the taxa are assigned to multiple ASVs, so the data is first aggregated
```{r}
# Aggregate the data by summing the Abundance values
aggregated_df <- myco.candida.df %>%
  group_by(name, Species, type, age, agegroup) %>%
  summarize(Abundance = sum(Abundance, na.rm = TRUE), .groups = 'drop')

# Label Function
labels <- c(Duodenum = "Duodenum", Jejunum = "Jejunum", Ileum = "Ileum", Caecum = "Caecum", ProxColon = "PC", DistalColon = "DC")

# Reorder regions
aggregated_df$type <- factor(aggregated_df$type, levels=c("Duodenum", "Jejunum", "Ileum", "Caecum", "ProxColon", "DistalColon"))

# Order ids by age (using a unique identifier to account for duplicates in df)
aggregated_df <- aggregated_df[order(aggregated_df$age, aggregated_df$name), ]
aggregated_df$name <- factor(aggregated_df$name, levels = unique(aggregated_df$name))


candida <- ggplot(aggregated_df, aes(name, Species, fill = Species, size = Abundance * 100)) +
  theme_light() +
  geom_point(alpha = 0.8, shape = 21) +
  scale_size(name = "Abundance", range = c(0, 6)) +
  facet_grid(~ type, space = "free", scales = "free", labeller = labeller(type = labels)) +
  theme(axis.text.x = element_text(angle = 90, size=12, hjust=0.8, vjust=0.2), 
        axis.text.y = element_text(size=12),
        axis.title = element_text(size = 16),
        axis.title.y = element_blank(),
        plot.title = element_text(face = "bold"),
        panel.grid.major = element_blank(), panel.grid.minor = element_blank(), 
        strip.background =element_rect(fill="grey"),
        strip.text = element_text(size = 12),  # Adjust the size of facet labels
        legend.text.align = 0, 
        legend.text = element_text(size=16),
        legend.title = element_text(size=18),
        legend.position = "right") +
  scale_fill_manual(values = cb_palette) +
  xlab("Sample ID") +
  guides(fill = "none") 

candida
```

* Make age group annotation plot
```{r}
# Rename Mid as Adult
aggregated_df$agegroup<-gsub("Mid","Adult",as.character(aggregated_df$agegroup))

# Reorder agegroup
aggregated_df$agegroup <- factor(aggregated_df$agegroup, levels = c("Young", "Adult", "Aged"))

# Create the age group annotations plot
age_group_annotations <- ggplot(aggregated_df, aes(x = name, fill = agegroup)) +
  geom_tile(aes(y = -5, height = 3)) +  # Adjust height to fit beneath the main plot
  scale_fill_manual(values = c("#f4e285", "#f4a259", "#8cb369")) +  # Customize colors as needed
  facet_grid(~ type, space = "free", scales = "free") +
  theme_void() +
  theme(legend.position = "bottom",
        strip.text = element_blank(),
        axis.text.x = element_blank(),
        axis.title = element_blank(),
        panel.spacing = unit(0.25, "lines"),
        legend.title = element_text(size = 16),
        legend.text = element_text(size = 14)) +
  guides(fill=guide_legend(title="Age Group"))

# Combine the two plots using patchwork
ca_combined_plot <- candida / age_group_annotations + plot_layout(heights = c(10, 1))
# Adjust heights as needed
ca_combined_plot
```

* Export graph
```{r}
tiff(file.path(folder_path, "CandidaSp.tiff"), units="in", width=12, height=5, res=300)
plot(ca_combined_plot)
dev.off()
```
